The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-01-27 paper/sections/methods.tex

%A pointform description of the general idea
%\begin{itemize}
%	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
%	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface. This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
%	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators. This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
%	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5). From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
%	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
%	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
%	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient. We augment our kinetic and potential energies with a stability term (ref VEM). This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
%	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction. We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
%	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion. Our generalized coordinates give us direct updates on the control points of the NURBS. Currently we are using linearly-implicit Euler for time integration.
%	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary. To compute these points we use an approach similar to the raycasting quadrature but in 2D. In the paremtric domain for a T-NURBS we sample uniforly in the V direction and shoot rays along the U axis. At each region of intersection, we sample more values along the ray. The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
%\end{itemize}
\section{Overview}
The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
(density, constitutive model, model parameters etc). 
The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
The output is an elastodynamic simulation. 
SEM consists of preprocessing and runtime simulation phases~\refalg{sem} and at no point do we need to generate a volumetric mesh of the input. 
In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
Finally we construct local shape matching operators and construct the mass matrix for our problem.

@@

\begin{algorithm}[h]
    \label{alg:sem}
	\caption{Shape Matching Element Algorithm}\label{sem_algorithm}
	\begin{algorithmic}[1]
		\Procedure{PrepareSim}{$model$}
		\State{// Get initial DOF and sample boundary  \Comment{Section \ref{sec:geometric_model}}}
		\State $\vc{q} \gets \text{GetControlPoints}(model)$
		\State $\samp{\Jnurbs}, \wt \gets \text{SampleNURBS}(model)$
		\State $\vc{x^0} \gets \wt \samp{\Jnurbs} \vc{q}$
        \Statex

@@

		\State{// Form mass matrix and error hessian, respectively}
    	\State $\MM \gets \text{MassMatrix}(\st{X}, \XBs, \st{W},\Pi)$    \Comment{Section \ref{ssec:mass_matrix}}
     	\State $\ME \gets \text{ErrorMatrix}(\vc{x^0}, \XBs, \st{W},\Pi)$  \Comment{Section \ref{ssec:error_energy}}
%     	\myworries{todo: add volume}
%     	\While {$simulating$}
%     		\State{$\Pcox \gets \Pi\mathbf{q}$}
%     		\ForAll{$i \gets 1$ to $|\st{X}|$}
%     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
%				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
%				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
%				
%				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
%				 \frac{\partial \mathbf{F}}{\partial c}^T 
%				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
%				 \frac{\partial \mathbf{F}}{\partial c}
%				 $}
%  			\EndFor
%  			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
%  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
%  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
%  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
%  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
%  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
%  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
%     	\EndWhile
	\EndProcedure
\end{algorithmic}
\caption{Shape Matching Element Algorithm}
\end{algorithm}

@@

The three-dimensional position, $\defX$, of any point on a NURBS surface can be written as 
\begin{equation}
\label{eqn:nurbs_srf}
\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^n \phi_{ij}\left(u,v\right)\vc{q}_{ij},
    %\mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %\text{,}
\end{equation}
where $u$ and $v$ $\in \real$ are the coordinates of the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
basis functions given by 
\begin{equation*}
    \phi_{ij}\left(u,v\right) = \frac{w_{i,j}B_{ik}(u)B_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} w_{ij} B_{ik}(u)B_{jl}(v)},
\end{equation*} where $B_{ik}$ (resp $B_{jl}$) is the $i^{th}$ ($j^{th}$) b-spline basis of order $k$ ($l$).

Exploiting linearity with respect to the control points  we can express this mapping as 
\begin{equation}
    \defX\left(u,v\right) = \underbrace{\begin{pmatrix} \phi_{11}\left(u,v\right)\ident & \dots & \phi_{nn}\left(u,v\right)\ident \end{pmatrix}}_{\Jnurbs\left(u,v\right)}
    \underbrace{\begin{pmatrix} \vc{q}_{11} \\ \vdots \\ \vc{q}_{nn}\end{pmatrix}}_{\vc{q}},
\end{equation} where $\Jnurbs\left(u,v\right)$ is the NURBS Jacobian, $\ident$ is the $3\times3$ identity matrix and $\vc{q}$ is the vector of generalized coordinates~\dave{cite Lanczos}.
The Shape Matching Element Method will follow from this kinematic description.

@@

>>>>

%A pointform description of the general idea
%\begin{itemize}
%	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
%	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface. This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
%	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators. This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
%	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5). From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
%	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
%	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
%	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient. We augment our kinetic and potential energies with a stability term (ref VEM). This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
%	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction. We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
%	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion. Our generalized coordinates give us direct updates on the control points of the NURBS. Currently we are using linearly-implicit Euler for time integration.
%	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary. To compute these points we use an approach similar to the raycasting quadrature but in 2D. In the paremtric domain for a T-NURBS we sample uniforly in the V direction and shoot rays along the U axis. At each region of intersection, we sample more values along the ray. The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
%\end{itemize}
\section{Overview}
The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
(density, constitutive model, model parameters etc). 
The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
The output is an elastodynamic simulation. 
SEM consists of preprocessing and runtime simulation phases (~\refalg{sem_preprocess}, ~\refalg{sem_simulation}) and at no point do we need to generate a volumetric mesh of the input. 
In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
Finally we construct local shape matching operators and construct the mass matrix for our problem.

@@

\begin{algorithm}[h]
	\begin{algorithmic}[1]
		\Procedure{Preprocess}{$model$}
		\State{// Get initial DOF and sample boundary  \Comment{Section \ref{sec:geometric_model}}}
		\State $\vc{q} \gets \text{GetControlPoints}(model)$
		\State $\samp{\Jnurbs}, \wt \gets \text{SampleNURBS}(model)$
		\State $\vc{x^0} \gets \wt \samp{\Jnurbs} \vc{q}$
        \Statex

@@

		\State{// Form mass matrix and error hessian, respectively}
    	\State $\MM \gets \text{MassMatrix}(\st{X}, \XBs, \st{W},\Pi)$    \Comment{Section \ref{ssec:mass_matrix}}
     	\State $\ME \gets \text{ErrorMatrix}(\vc{x^0}, \XBs, \st{W},\Pi)$  \Comment{Section \ref{ssec:error_energy}}
	\EndProcedure
\end{algorithmic}
\caption{Shape Matching Element Preprocessing}
\label{alg:sem_preprocess}
\end{algorithm}

@@

The three-dimensional position, $\defX$, of any point on a NURBS surface can be written as 
\begin{equation}
\label{eqn:nurbs_srf}
\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^m \phi_{ij}\left(u,v\right)\vc{q}_{ij},
    %\mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %\text{,}
\end{equation}
where $u$ and $v$ $\in \real$ are the coordinates in the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
basis functions given by 
\begin{equation*}
    \phi_{ij}\left(u,v\right) = \frac{w_{i,j}B_{ik}(u)B_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} w_{ij} B_{ik}(u)B_{jl}(v)},
\end{equation*} where $B_{ik}$ (resp. $B_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$).

Exploiting linearity with respect to the control points, we can express this mapping as
\begin{equation}
    \defX\left(u,v\right) = \underbrace{\begin{pmatrix} \phi_{11}\left(u,v\right)\ident & \dots & \phi_{nn}\left(u,v\right)\ident \end{pmatrix}}_{\Jnurbs\left(u,v\right)}
    \underbrace{\begin{pmatrix} \vc{q}_{11} \\ \vdots \\ \vc{q}_{nm}\end{pmatrix}}_{\vc{q}},
\end{equation} where $\Jnurbs\left(u,v\right)$ is the NURBS Jacobian, $\ident$ is the $3\times3$ identity matrix and $\vc{q}$ is the vector of generalized coordinates~\dave{cite Lanczos}.
The Shape Matching Element Method will follow from this kinematic description.

@@

\begin{algorithm}[h]
	\begin{algorithmic}[1]
		\Procedure{Preprocess}{$model$}
     	\myworries{todo: add volume}
     	\While {$simulating$}
     		\State{$\Pcox \gets \Pi\mathbf{q}$}
     		\ForAll{$i \gets 1$ to $|\st{X}|$}
     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}

				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
				 \frac{\partial \mathbf{F}}{\partial c}^T 
				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
				 \frac{\partial \mathbf{F}}{\partial c}
				 $}
  			\EndFor
  			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
     	\EndWhile
	\EndProcedure
\end{algorithmic}
\caption{Shape Matching Element Simulation Loop}
\label{alg:sem_simulation}
\end{algorithm}
